\bf Title here

### 2020-06-23 paper/paper.Rmd
write up the skeleton of diagnostic plot in theory and rearrange examples

## Visual diagnostics 
<!-- Visual diagnostics can be real-time or post-run. Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.   This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics. -->

@@

@@

@@

@@

@@

## Explore scalar parameters

### Explore searching points

@@

### Explore interpolating points

Plotting the interpolating points as time series data allows us to diagnose characteristics of different configurations and index functions. Here we present two examples of using  plots to diagnose tour algorithms and different index functions.

```{r interpolation}
# points is always included in the newest version
# p1 <- holes_2d_better %>% 
#   filter(info == "interpolation") %>% 
#   ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
#   geom_point() + 
#   geom_line(aes(group = 1))
# 
# p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
# 
# (p1 | p2)  &
#   theme(legend.position = "bottom")

@@

### Adding more variables to the mapping

@@

## Explore matrix parameter

@@

@@

@@

# Animated diagnostic plots

```{r pca-animated, echo = TRUE, eval = FALSE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) + 
  theme(legend.position = "bottom")
```

>>>>

## Origin idea
<!-- Visual diagnostics can be real-time or post-run. Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.   This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics. -->

@@

An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore... [the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part. ]

@@

## Visual diagnostics

Follow the research question we raised earlier, the purpose of visual diagnostics is to understand

- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.

- How does the searching space look like, that is, geometrically, where are the points located in the space.

Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function. The second question can be addressed via visualising the rotating high dimensional space. This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time. The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically. Thus we can do a similar rotation on the screen to precieve even higher dimension. [there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper]. Thus animated visualisation is needed to preceive the optimisation path in the searching space.

The plot design of visual diagnostics depends on the characteristics of the variables to plot. If the searching points are of interest [point geom is not good; need summarise. While updating points are more manageable -> point geom]. However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point. [stratefy may help? - may need more work here.]

Exploring the value of an objective function is cna be a simple task. Because the iterator has a time series ordering feature, it goes onto the x-axis as convention. The value of hte objective function goes to the y-axis. Other parameter of interest can be represented using color to present more information. [This is the case for our purpose with method, but when the level goes higher or the parameter is quantitative, color may not be a good option. ]

An option to explore of the searching space is to explore a reduced projection 2D space via principle component analysis or other dimensional reduction methods. In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution. [need to think further about how different type of variables can be shown]. While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation. Geozoo is an option. It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.

@@

From Section \ref{tour}, we presents a comprehensive case study on how to use the visual diagnostics to explore the optimisation in a specific problem: projection pursuit guided tour. Section \ref{tour} aims to provide non-experts with an overview of the problem content and the existing optimisation procedures used in projection pursuit guided tour. For those who are already familiar with the technique, this section can be skipped.

@@

In this section, we present the simulated data along with how the global object can be applied in guided tour.

@@

Below we will present several examples to diagnose different aspect of the tour optimisation. We will first provide examples on producing 1) static plots to explore the value of objective function and 2) animated plots to explore the searching space and then a more sophisticated example that combines both to understand how to optimise a noisy and complex index function. In both section, we will first provide a toy example that is easy to grasp and one or two examples that help us understand the optimisation algorithm and parameter choice better.

## Explore the value of objective function {#static}

### Toy example: exploring searching points

@@

### A more complex example: adding an interruption in the interpolation stage

```{r interpolation}
#points is always included in the newest version
# p1 <- holes_2d_better %>%
#   filter(info == "interpolation") %>%
#   ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
#   geom_point() +
#   geom_line(aes(group = 1))
# 
# p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
# 
# (p1 | p2)  &
#   theme(legend.position = "bottom")
# 
```

@@

### Another example: polishing the final projection basis

@@

## Explore searching space {#animated}

@@

### A toy example: understand different stage of search_geodesic

@@

### A more complex example: Choosing the initial value for polishing parameter

@@

<!-- # Animated diagnostic plots -->

<!-- ```{r pca-animated, echo = TRUE, eval = FALSE} -->
<!-- holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) +  -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->

## A comprehensive example of diagnosing a noisy index function

### 2020-06-23 paper/paper.tex
write up the skeleton of diagnostic plot in theory and rearrange examples

@@

@@

\hypertarget{explore-scalar-parameters}{%
\subsection{Explore scalar parameters}\label{explore-scalar-parameters}}

\hypertarget{explore-searching-points}{%
\subsubsection{Explore searching
points}\label{explore-searching-points}}

@@

\hypertarget{explore-interpolating-points}{%
\subsubsection{Explore interpolating
points}\label{explore-interpolating-points}}

Plotting the interpolating points as time series data allows us to
diagnose characteristics of different configurations and index
functions. Here we present two examples of using plots to diagnose tour
algorithms and different index functions.

@@

\hypertarget{adding-more-variables-to-the-mapping}{%
\subsubsection{Adding more variables to the
mapping}\label{adding-more-variables-to-the-mapping}}

@@

\hypertarget{explore-matrix-parameter}{%
\subsection{Explore matrix parameter}\label{explore-matrix-parameter}}

@@

@@

@@

\hypertarget{animated-diagnostic-plots}{%
\section{Animated diagnostic plots}\label{animated-diagnostic-plots}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =}\NormalTok{ info) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

>>>>

\hypertarget{origin-idea}{%
\subsection{Origin idea}\label{origin-idea}}

@@

An optimisation procedure can easily generate hundreds or thousands
sampling points and the question becomes what are the points we want to
explore\ldots{} {[}the classification of searching or updating points
goes here and people should be clear what part of the points goes to the
plot - this is the data part not the plot part. {]}

Follow the research question we raised earlier, the purpose of visual
diagnostics is to understand

\begin{itemize}
\item
  Whether the procedure has provided an optimum and how the value of the
  objective function changes through the optimisation.
\item
  How does the searching space look like, that is, geometrically, where
  are the points located in the space.
\end{itemize}

Overall, the first question can be answered using a static plot with
x-axis showing the progression of the optimisation and y-axis showing
the value of the objective function. The second question can be
addressed via visualising the rotating high dimensional space. This
seems to be a hard task to see beyond 3D, but the fact is that we never
really see all the six faces of a cube at one time. The way humans
understand the dimension of a 3D object is either through rotating the
physical 3D object or using shade and line type to annotate the 3D
object in a paper or electronically. Thus we can do a similar rotation
on the screen to precieve even higher dimension. {[}there is likely to
be a learning curve - remember when learning geometry in school, it
takes a while to get used to see 3D cube on a paper{]}. Thus animated
visualisation is needed to preceive the optimisation path in the
searching space.

The plot design of visual diagnostics depends on the characteristics of
the variables to plot. If the searching points are of interest {[}point
geom is not good; need summarise. While updating points are more
manageable -\textgreater{} point geom{]}. However, one must realise that
with hundreds or thousands of searching points, exploring the sample
space in animation could be slow and this is because of the time it
takes to render hundreds point. {[}stratefy may help? - may need more
work here.{]}

Exploring the value of an objective function is cna be a simple task.
Because the iterator has a time series ordering feature, it goes onto
the x-axis as convention. The value of hte objective function goes to
the y-axis. Other parameter of interest can be represented using color
to present more information. {[}This is the case for our purpose with
method, but when the level goes higher or the parameter is quantitative,
color may not be a good option. {]}

An option to explore of the searching space is to explore a reduced
projection 2D space via principle component analysis or other
dimensional reduction methods. In this way, the first two principle
components will take up the x and y axis and all the other information
will be mapped using other aesthetic attribution. {[}need to think
further about how different type of variables can be shown{]}. While
explore the reduced space is an initial attempt to understand the
searching space, there are existing technology for rotating a higher
dimensional space for visualisation. Geozoo is an option. It generates
random points on the high dimensional space and we can overlay it with
the points on the optimisation path to visualise the spread of it on the
high-D sphere.

@@

Below we will present several examples to diagnose different aspect of
the tour optimisation. We will first provide examples on producing 1)
static plots to explore the value of objective function and 2) animated
plots to explore the searching space and then a more sophisticated
example that combines both to understand how to optimise a noisy and
complex index function. In both section, we will first provide a toy
example that is easy to grasp and one or two examples that help us
understand the optimisation algorithm and parameter choice better.

\hypertarget{static}{%
\subsection{Explore the value of objective function}\label{static}}

\hypertarget{toy-example-exploring-searching-points}{%
\subsubsection{Toy example: exploring searching
points}\label{toy-example-exploring-searching-points}}

@@

\hypertarget{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}{%
\subsubsection{A more complex example: adding an interruption in the
interpolation
stage}\label{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}}

@@

\hypertarget{another-example-polishing-the-final-projection-basis}{%
\subsubsection{Another example: polishing the final projection
basis}\label{another-example-polishing-the-final-projection-basis}}

@@

\hypertarget{animated}{%
\subsection{Explore searching space}\label{animated}}

@@

\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
\subsubsection{A toy example: understand different stage of
search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}

@@

\hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
\subsubsection{A more complex example: Choosing the initial value for
polishing
parameter}\label{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}}

@@

\hypertarget{a-comprehensive-example-of-diagnosing-noisy-index-function}{%
\subsection{A comprehensive example of diagnosing noisy index
function}\label{a-comprehensive-example-of-diagnosing-noisy-index-function}}
